---
title: "STATS506 Problem Set #04"
author: "Fan Zhang"
date: "10/19/2023"
format:
    html:
        embed-resources: true 
        code-line-numbers: true
        # code-fold: true
editor: 
    render-on-save: false 
---

GitHub Repository: [https://github.com/cajole-fff/UMich-STATS506-FanZhang/](https://github.com/cajole-fff/UMich-STATS506-FanZhang/)

## Problem 1 - Tidyverse 

Use the `tidyverse` for this problem. In particular, use piping and `dplyr` as much as you are able. **Note:** Use of any deprecated functions will result in a point loss. 

Install and load the package `nycflights13`.


```{r} 
# Load the packages
library(nycflights13)
library(tidyverse)
```

**a. Generate a table (which can just be a nicely printed tibble) reporting the mean and median departure delay per airport. Generate a second table (which again can be a nicely printed tibble) reporting the mean and median arrival delay per airport. Exclude any destination with under 10 flights. Do this exclusion through code, not manually.**

**Additionally,**

- Order both tables in descending mean delay.
- Both tables should use the airport names not the airport codes.
- Both tables should print all rows. 

```{r} 
# Generate table for mean and median departure delay per airport
flights %>%
    inner_join(airports, by = c("origin" = "faa")) %>%
    group_by(name) %>%
    summarise(mean_dep_delay = mean(dep_delay, na.rm = TRUE),median_dep_delay = median(dep_delay, na.rm = TRUE)) %>%
    arrange(desc(mean_dep_delay)) %>%
    print()
```

```{r} 
# Generate table for mean and median arrival delay per airport
flights %>%
    inner_join(airports, by = c("dest" = "faa")) %>%
    group_by(name) %>%
    filter(n() >= 10) %>%
    summarise(mean_arr_delay = mean(arr_delay, na.rm = TRUE),median_arr_delay = median(arr_delay, na.rm = TRUE)) %>%
    arrange(desc(mean_arr_delay)) %>%
    print()
```

**b. How many flights did the aircraft model with the fastest average speed take? Produce a tibble with 1 row, and entires for the model, average speed (in MPH) and number of flights.**

```{r}
# Generate table for the aircraft model with the fastest average speed
flights %>%
    inner_join(planes, by = c("tailnum")) %>%
    group_by(model) %>%
    summarise(avg_speed = mean(speed, na.rm = TRUE), num_flights = n()) %>%
    filter(num_flights > 0) %>%
    arrange(desc(avg_speed)) %>%
    head(1) %>%
    print()
```

## Problem 2 - `get_temp()` 

Use the `tidyverse` for this problem. In particular, use piping and `dplyr` as much as you are able. **Note:** Use of any deprecated functions will result in a point loss. 

Load the Chicago NNMAPS data we used in the visualization lectures. Write a function `get_temp()` that allows a user to request the average temperature for a given month. The arguments should be: 

- `month`: Month, either a numeric 1-12 or a string.
- `year`: A numeric year.
- `data`: The data set to obtain data from.
- `celsius`: Logically indicating whether the results should be in celsius. Default `FALSE`.
- `average_fn`: A function with which to compute the mean. Default is `mean`. 

The output should be a numeric vector of length 1. The code inside the function should, as with the rest of this problem, use `tidyverse`. Be sure to sanitize the input. 

Prove your code works by evaluating the following. Your code should produce the result, or a reasonable error message. 

```{r, eval = FALSE}
get_temp("Apr", 1999, data = nnmaps)
get_temp("Apr", 1999, data = nnmaps, celsius = TRUE)
get_temp(10, 1998, data = nnmaps, average_fn = median)
get_temp(13, 1998, data = nnmaps)
get_temp(2, 2005, data = nnmaps)
get_temp("November", 1999, data =nnmaps, celsius = TRUE,
        average_fn = function(x) {
            x %>% sort -> x
            x[2:(length(x) - 1)] %>% mean %>% return
        })
```

```{r} 
# Load the package and data 
library(tidyverse)
nnmaps <- read.csv("./chicago-nmmaps.csv")
head(nnmaps)
```

```{r}
# Define the function
get_temp <- function(month, 
                    year, 
                    data, 
                    celsius = FALSE, 
                    average_fn = mean) {
    # Sanitize the input
    months_abb <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
    months_num <- c(1:12)
    months_name <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

    if (month %in% months_abb) {
        month <- match(month, months_abb)
    } else if (month %in% months_name) {
        month <- match(month, months_name)
    }
    else if (month %in% months_num) {
        month <- as.numeric(month)
    } else {
        cat("Invalid month input", month, ". Should be an integer between 1 and 12 (for example, 1)or a string of month name (for example, January) or abbreviation (for example, Jan). Not case sensitive.\n")
        return 
    }

    year <- as.integer(year)

    if (!is.integer(year) || year < 1997 || year > 2000) {
        cat("Invalid year input", year, ". Should be an integer between 1997 and 2000 (inclusive).\n")
        return 
    }
    if (!is.data.frame(data)) {
        cat("Invalid data input", data, ". Should be a data frame.\n")
        return 
    }
    if (!is.logical(celsius)) {
        cat("Input celsius input", celsius, ". Should be a logical value TRUE or FALSE.\n")
        return 
    }
    if (!is.function(average_fn)) {
        cat("Input average_fn input", average_fn, ". Should be a function.\n")
        return 
    }
    month_tmp <- month
    year_tmp <- year 
    # print(month_tmp, year_tmp)
    # Compute the average temperature
    data %>%
        filter(month_numeric == month_tmp & year == year_tmp) %>%
        summarise(avg_temp = average_fn(temp)) %>%
        mutate(avg_temp = ifelse(celsius == TRUE, (avg_temp - 32) * 5 / 9, avg_temp)) %>%
        pull(avg_temp) %>%

    # Return a numeric vector of length 1
    return
}
```

```{r}
# Demo
get_temp("Apr", 1999, data = nnmaps)
get_temp("Apr", 1999, data = nnmaps, celsius = TRUE)
get_temp(10, 1998, data = nnmaps, average_fn = median)
get_temp(13, 1998, data = nnmaps)
get_temp(2, 2005, data = nnmaps)
get_temp("November", 1999, data =nnmaps, celsius = TRUE,
        average_fn = function(x) {
        x %>% sort -> x
        x[2:(length(x) - 1)] %>% mean %>% return
        })
```


## Problem 3 - SAS 
